CRUSE to Safe Cycling in Ireland

### 2023-09-21 paper.qmd
Merge GOT input

<!-- \[Gary, Suzanne, others input here please.\] -->

It was in this context that Transport Infrastructure Ireland (TII) commissioned the Cycle Route Uptake and Scenario Estimation (CRUSE) project.
Based on the Propensity to Cycle Tool (PCT) for England and Wales [@lovelace2017], the CRUSE tool was developed to provide evidence on current cycling levels and future cycling potential nationwide across Ireland.

@@

The methods used to generate the evidence presented in the CRUSE tool for Ireland build on the Propensity to Cycle Tool (PCT), which was originally funded by the UK's Department for Transport and developed by a multi-university team.
An important feature of the PCT is that it is open source and publicly available (at [www.pct.bike](https://www.pct.bike/)), allowing its use by all stakeholders in the transport planning process [@lovelace2017].
The PCT approach has had major policy impacts, as outlined in Research Excellence Framework (REF) impact case studies, which demonstrate that the tool "revolutionised strategic cycle planning in England and Wales"[^2] by overcoming the barriers to cycling investment imposed by lack of evidence on cycling potential[^3]
.

[^2]: See REF Impact Case Study "Cycle network policy, planning and investment transformed by the Propensity to Cycle Tool" at [results2021.ref.ac.uk](https://results2021.ref.ac.uk/impact/847d1191-7f25-46ba-a399-b481125edc8f) submitted by the University of Leeds.

[^3]: See the REF Impact Case Study "Creating Step Changes in Cycling Policy and Infrastructure Planning across the UK" at [results2021.ref.ac.uk](https://results2021.ref.ac.uk/impact/4BBF3436-FD10-4C75-9791-F5E98AB4411B) submitted by the University of Westminster.

@@

The PCT approach has been applied in other countries, including Ireland (the topic of this paper), Scotland, and Portugal.
In Portugal the 'biclaR' project, based on methods underlying the PCT, has been developed and deployed for the Lisbon metro region.
The resulting evidence is available in an interactive web application hosted at [biclar.tmlmobilidade.pt](https://biclar.tmlmobilidade.pt) [@felix2023].
biclaR includes estimates of impacts, using the World Health Organisation (WHO) 'HEAT for Cycling' tool[^4] and an 'intermodality' scenario that combines cycling with currently available public transit options based on General Transit Feed Specification (GTFS) data.

[^4]: See the Health economic assessment tool for walking and cycling at [www.heatwalkingcycling.org](https://www.heatwalkingcycling.org/#homepage).

>>>>

A proactive approach to the development of cycling policy and
infrastructure is taking place in Ireland.
The main framework
underpinning these efforts is the Climate Action Plan[^2] and The National
Development Plan[^3]
. Strategic alignment is also outlined by the
Phoenix Park Transport and Mobility Options Study, which highlights public support for active travel interventions: \"83% of respondents supported enhanced walking and cycling facilities\"
.

[^2]: <https://www.gov.ie/en/publication/7bd8c-climate-action-plan-2023/>

[^3]: <https://www.gov.ie/en/publication/774e2-national-development-plan-2021-2030/>

At the regional level, the recently published Greater Dublin Area Transport Strategy[^4] reinforces these findings: \"nearly a quarter of adults cycle at least once a week in the Dublin Metropolitan Area\" with cycling in the Dublin area taking up to 60,000 cars off the road today.
Extrapolating this on a per population basis across Ireland, with around 40% of the population living in Dublin, this suggests that around 150,000 cars could be removed nationwide just by achieving Dubline levels of cycling in all counties (notwithstanding existing cycling trips and differences in trip distances).

[^4]: <https://www.nationaltransport.ie/planning-and-investment/strategic-planning/greater-dublin-area-transport-strategy/>

Further evidence of the importance of cycling in Ireland *already* is provided by the National Strategic Objective (NSO) from the National Development Plan, with €8.6 billion allocated to sustainable transport infrastructure including public transport and active travel interventions.
Cycle infrastructure will be developed in synchrony with the BusConnects project, an entire redesign of the bus network in Dublin and Cork.

It was in this context that Transport Infrastructure Ireland (TII) commissioned the Cycle Route Uptake and Scenario Estimation (CRUSE) project.
Based on the Propensity to Cycle Tool (PCT) for England and Wales, the CRUSE tool was developed to provide evidence on current cycling levels and future cycling potential nationwide across Ireland.
A key consideration for TII is to ensure both urban and rural Ireland were targeted for cycling infrastructure.
To this end, TII extensively reviewed activity on Irish Roads and how the road network was being used by commuters, cyclists, heavy goods vehicles and other road users across the Irish transport network.
The aim was for the tool to provide strong, national, systematic but locally-specific evidence to monitor cycling friendliness and safety in Ireland, to ensure strategic alignment with national, regional and local policies.
The open source and publicly available nature of the tool ensures more inclusive and evidence-based conversations around cycle network planning between all stakeholders in the transport planning process.
Other use cases in the Irish context includes monitoring the need for existing cycling infrastructure upgrades, undertaking appraisal, as well as safety and exposure information in line with the European Cycling Federation, Dutch Cycling Embassy recommendations and the European RISM Directive, for reporting collision rates of vulnerable road users by 2024.
These factors mean that the tool can be seen as an open access 'leverage point' in the planning system [@lovelace2020].

@@

The methods used to generate the evidence presented in the CRUSE tool for Ireland build on the Propensity to Cycle Tool (PCT), which was originally funded by the UK's Department for Transport and developed by a multi-university team.
An important feature of the PCT is that it is open source and publicly available (at [www.pct.bike](https://www.pct.bike/)), allowing its use by all stakeholders in the transport planning process [@lovelace2017].
The PCT approach has had major policy impacts, as outlined in Research Excellence Framework (REF) impact case studies, which demonstrate that the tool "revolutionised strategic cycle planning in England and Wales"[^5] by overcoming the barriers to cycling investment imposed by lack of evidence on cycling potential[^6]
.

[^5]: See REF Impact Case Study "Cycle network policy, planning and investment transformed by the Propensity to Cycle Tool" at [results2021.ref.ac.uk](https://results2021.ref.ac.uk/impact/847d1191-7f25-46ba-a399-b481125edc8f) submitted by the University of Leeds.

[^6]: See the REF Impact Case Study "Creating Step Changes in Cycling Policy and Infrastructure Planning across the UK" at [results2021.ref.ac.uk](https://results2021.ref.ac.uk/impact/4BBF3436-FD10-4C75-9791-F5E98AB4411B) submitted by the University of Westminster.

@@

The PCT approach has been applied in other countries, including Ireland (the topic of this paper), Scotland, and Portugal.
In Portugal the 'biclaR' project, based on methods underlying the PCT, has been developed and deployed for the Lisbon metro region.
The resulting evidence is available in an interactive web application hosted at [biclar.tmlmobilidade.pt](https://biclar.tmlmobilidade.pt) [@felix2023].
biclaR includes estimates of impacts, using the World Health Organisation (WHO) 'HEAT for Cycling' tool[^7] and an 'intermodality' scenario that combines cycling with currently available public transit options based on General Transit Feed Specification (GTFS) data.

[^7]: See the Health economic assessment tool for walking and cycling at [www.heatwalkingcycling.org](https://www.heatwalkingcycling.org/#homepage).
